Semi-Parametric Manifold Clustering

### 2021-11-03 curve-fitting-methods/curve-fitting-scratch.Rmd
minor updates

$$
g(t) = (1 - t)^2 p_0 + 2 t (1 - t) p_1 + t^2 p_2
$$

@@

Common methods for Bezier curve fitting do not require specification of $t_1, ..., t_n$. However, an ordering is required. The curve fitting method is then modified as follows:

1. Initialize $\hat{\theta}^{(0)}$ (e.g., randomly). 
2. Estimate each $\hat{t}_i^{(s)}$ by minimizing $L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2$.
3. Reorder $X$ according to $t_1, ..., t_n$.
4. Estimate $\hat{p}_0^{(s)}, \hat{p}_1^{(s)}, \hat{p}_2^{(s)}$ for the reordered $X$ using Bezier curve fitting.
5. Compute $\hat{\theta}^{(s+1)}$ from $\{\hat{p}_r^{(s)}\}$.
6. Repeat steps 2-5 until convergence.

>>>>

$$
g(t | p) = (1 - t)^2 p_0 + 2 t (1 - t) p_1 + t^2 p_2
$$

@@

Common methods for Bezier curve fitting do not require specification of $t_1, ..., t_n$. 
However, an ordering is required.

The curve fitting method is then modified as follows:

1. Initialize $\hat{p}^{(0)}$ (e.g., randomly). 
2. Estimate each $\hat{t}_i^{(s)}$ by minimizing $L(t_i, \hat{p}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{p}^{(s)})\|^2$.
3. Reorder $X$ according to $t_1, ..., t_n$.
4. Estimate $\hat{p}^{(s)}$ for the reordered $X$ using Bezier curve fitting.
5. Repeat steps 2-4 until convergence.

Then this becomes step 2 of the clustering method outlined above.

We can also estimate $\hat{p} = (T^\top T)^{-1} T^\top X$ where $T = \begin{bmatrix} t_1 & \cdots & t_R \end{bmatrix}$ and each $t_r = \binom{R}{r} (1-t)^{R-r} t^r$, where $R$ is the order of the Bezier curve, so using the method for estimating $t_1, ..., t_n$ as described before, we can iteratively solve for $\hat{p}$ in the same manner. If we use soft clustering (e.g., EM as opposed to $K$-means), we can also insert a diagonal matrix of weights $W_k$ to obtain

$$\hat{p}_k = (T^\top W_k T)^{-1} T^\top W_k X$$

where $X$ is the full data matrix, $T$ is the full model matrix, and $W_k = w_k$ with the entries of $w_k$ being the estimated probability of each $x_i$ being in cluster $k$.

### 2021-11-03 curve-fitting-methods/curve-fitting-scratch.tex
minor updates

\[
g(t) = (1 - t)^2 p_0 + 2 t (1 - t) p_1 + t^2 p_2
\]

@@

Common methods for Bezier curve fitting do not require specification of
\(t_1, ..., t_n\). However, an ordering is required. The curve fitting
method is then modified as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Initialize \(\hat{\theta}^{(0)}\) (e.g., randomly).
\item
  Estimate each \(\hat{t}_i^{(s)}\) by minimizing
  \(L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2\).
\item
  Reorder \(X\) according to \(t_1, ..., t_n\).
\item
  Estimate \(\hat{p}_0^{(s)}, \hat{p}_1^{(s)}, \hat{p}_2^{(s)}\) for the
  reordered \(X\) using Bezier curve fitting.
\item
  Compute \(\hat{\theta}^{(s+1)}\) from \(\{\hat{p}_r^{(s)}\}\).
\item
  Repeat steps 2-5 until convergence.
\end{enumerate}

>>>>

\[
g(t | p) = (1 - t)^2 p_0 + 2 t (1 - t) p_1 + t^2 p_2
\]

@@

Common methods for Bezier curve fitting do not require specification of
\(t_1, ..., t_n\). However, an ordering is required.

The curve fitting method is then modified as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Initialize \(\hat{p}^{(0)}\) (e.g., randomly).
\item
  Estimate each \(\hat{t}_i^{(s)}\) by minimizing
  \(L(t_i, \hat{p}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{p}^{(s)})\|^2\).
\item
  Reorder \(X\) according to \(t_1, ..., t_n\).
\item
  Estimate \(\hat{p}^{(s)}\) for the reordered \(X\) using Bezier curve
  fitting.
\item
  Repeat steps 2-4 until convergence.
\end{enumerate}

Then this becomes step 2 of the clustering method outlined above.

We can also estimate \(\hat{p} = (T^\top T)^{-1} T^\top X\) where
\(T = \begin{bmatrix} t_1 & \cdots & t_R \end{bmatrix}\) and each
\(t_r = \binom{R}{r} (1-t)^{R-r} t^r\), where \(R\) is the order of the
Bezier curve, so using the method for estimating \(t_1, ..., t_n\) as
described before, we can iteratively solve for \(\hat{p}\) in the same
manner. If we use soft clustering (e.g., EM as opposed to \(K\)-means),
we can also insert a diagonal matrix of weights \(W_k\) to obtain

\[\hat{p}_k = (T^\top W_k T)^{-1} T^\top W_k X\]

where \(X\) is the full data matrix, \(T\) is the full model matrix, and
\(W_k = w_k\) with the entries of \(w_k\) being the estimated
probability of each \(x_i\) being in cluster \(k\).
